---
title: "team"
author: "Brandon Xicon"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading Data: Cyclone Football 
```{r}
library(readxl)
Defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
Offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
Bio <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
str(Defense)#checks if it worked 

```



## Part one: Cleaning data
The three Excel sheets are loaded into R but they are not useful to us before some cleaning. Manipulate the three imported data frames, so that the cleaned data satisfy the following:

1. The player names and opponent names are factors.
2. All offensive and defensive statistics are numerical. The `Weight` column in Biography is also numeric. (Hint: look at `?mutate_at` if you want to save writing some code). 
3.Change the 'Height' column in biography into numeric (Hint: Use a unit that works well)
4. By cleaning the three data frames in R, at the end of this step you should have the following three clean data frames: `defClean`, `offClean`, and `bioClean`. Print the structures of these data frames. **Make sure your data types make sense for each variable.**
```{r}
#install and load needed packages
if (!require("tidyverse")) {
install.packages("tidyverse")
}
if (!require("dplyr")) {
install.packages("dplyr")
}
if (!require("ggplot2")) {
install.packages("ggplot2")
}

library(dplyr)
library(tidyverse)
library(ggplot2)

```


```{r}
#player names and opponent names are factors
Bio$Name <- as.factor(Bio$Name)
Defense$Name <- as.factor(Defense$Name)
Offense$Name <- as.factor(Offense$Name)
Defense$Opponent_Opponent <- as.factor(Defense$Opponent_Opponent)
Offense$Opponent_Opponent <- as.factor(Offense$Opponent_Opponent)

```


```{r}

#offensive and defensive statistics are numerical. 'Weight' column in biography is also numeric.
Defense <- mutate_at(Defense, 3:11, as.numeric)
Offense <- mutate_at(Offense, 3:12, as.numeric)
Bio$Weight <- as.numeric(Bio$Weight)
defClean <- Defense
offClean <- Offense
```

```{r}
#Change Height column in biography into numeric
#TODO my "%>%" wasnt recognized as a function for some reason
bioClean <- Bio %>% separate(Height, c("Feet", "Inches"))
bioClean$Feet = as.numeric(bioClean$Feet)
bioClean$Inches = as.numeric(bioClean$Inches)
bioClean$Height = (bioClean$Feet *12)+bioClean$Inches
bioClean = select(bioClean,-matches("Feet"))
bioClean = select(bioClean,-matches("Inches"))
bioClean <- bioClean[,c(1,2,7,3,4,5,6)] 
head(bioClean)
```

 ## Part two: Tidying (1)
For the following questions, work on the cleaned data frames defClean, offClean, and bioClean.

1. Reformat the defClean data frame into a tidy format using pivot_longer. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.

2. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?

3.Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense.
```{r}
#Reformatting the defClean data frame into a tidy format also Changed Statistic to a factor for ease of use in the future
defClean <- defClean %>% 
  pivot_longer(cols = c("Tackles_Solo","Tackles_Sack","Tackles_ASST", "Tackles_TFL","Turnover_FF", "Turnover_FR","Turnover_INT", "Pass_QBH","Pass_PB" ),
    names_to = "Statistic",
    values_to = "Value") 
defClean$Statistic <- as.factor(defClean$Statistic)

str(defClean)

```

```{r}
#Creating a Histogram For each Defensice Statistic

ggplot(defClean, aes(x=Value)) +
  geom_histogram() +
  facet_wrap(defClean$Statistic)
```
Looking at the histogram we see that almost all Stats have 0 as the majority howvever any Stat involing a turnover or sack is much more rare than other stats (IE the majority of players have 0 and almost no player has more than 1 in a game)

```{r}
#Did ISU have a better defense against West Virgina or Kanas State?
dat1 <- defClean
dat1 <- dat1 %>% 
  filter(Opponent_Opponent == "West Virginia" | Opponent_Opponent == "Kansas State") %>%
  filter(Statistic == "Tackles_Solo") %>%
  pivot_wider(names_from = Opponent_Opponent, values_from = Value) %>%
  mutate( WV = `West Virginia` ) %>%
  mutate(KS = `Kansas State`)


ggplot(dat1, aes(x=WV, y=KS)) + 
  geom_point(position = "jitter")

```

As you can see from the plot ISU had a better defense against Kansas State than against West Virginia, as more players had more solo tackles

## Part two: Tidying (2)
1. Separate the Hometown column into two columns, namely the city and the state. Print the first few rows of these two columns. (Hint: look at the sep= argument of separate)

2.How many players are there from each state?

```{r}
#Seperateing the Hometown Column into city and State
bioClean <- bioClean %>%
  separate(Hometown,into = c("City", "State"), sep= ',')
head(bioClean)
```

```{r}
#HOw many Players in each State
dat2 <- bioClean
bioClean %>% 
  count(State, sort = TRUE)
  
```

## Part Three: Joining data frames
1. Find where do the offensive players come from by joining the offensive data and player biography data appropriately.

2. Is there a relationship between the weight of a player and the receiving yards (Receiving_YDS)?

3. How do the performance of quarterback Brock Purdy relate to the performance of the ISU team?

4. Explore more. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.
