---
title: "team"
author: "Brandon Xicon"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading Data: Cyclone Football 
```{r}
library(readxl)
Defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
Offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
Bio <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
str(defense)#checks if it worked 

```



## Part one: Cleaning data
The three Excel sheets are loaded into R but they are not useful to us before some cleaning. Manipulate the three imported data frames, so that the cleaned data satisfy the following:

1. The player names and opponent names are factors.
2. All offensive and defensive statistics are numerical. The `Weight` column in Biography is also numeric. (Hint: look at `?mutate_at` if you want to save writing some code). 
3.Change the 'Height' column in biography into numeric (Hint: Use a unit that works well)
4. By cleaning the three data frames in R, at the end of this step you should have the following three clean data frames: `defClean`, `offClean`, and `bioClean`. Print the structures of these data frames. **Make sure your data types make sense for each variable.**
```{r}
#install and load needed packages
if (!require("tidyverse")) {
install.packages("tidyverse")
}
if (!require("dplyr")) {
install.packages("dplyr")
}

library(dplyr)
library(tidyverse)

```


```{r}
#player names and opponent names are factors
Bio$Name <- as.factor(Bio$Name)
Defense$Name <- as.factor(Defense$Name)
Offense$Name <- as.factor(Offense$Name)
Defense$Opponent_Opponent <- as.factor(Defense$Opponent_Opponent)
Offense$Opponent_Opponent <- as.factor(Offense$Opponent_Opponent)

```


```{r}

#offensive and defensive statistics are numerical. 'Weight' column in biography is also numeric.
Defense <- mutate_at(Defense, 3:11, as.numeric)
Offense <- mutate_at(Offense, 3:12, as.numeric)
bio$Weight <- as.numeric(Bio$Weight)
defClean <- Defense
offClean <- Offense
```

```{r}
#Change Height column in biography into numeric
#TODO my "%>%" wasnt recognized as a function for some reason
bioClean <- Bio %>% separate(Height, c("Feet", "Inches"))
bioClean$Feet = as.numeric(bioClean$Feet)
bioClean$Inches = as.numeric(bioClean$Inches)
bioClean$Height = (bioClean$Feet *12)+bioClean$Inches
bioClean = select(bioClean,-matches("Feet"))
bioClean = select(bioClean,-matches("Inches"))
bioClean <- bioClean[,c(1,2,7,3,4,5,6)] 
bioClean
```

 ## Part two: Tidying (1)
For the following questions, work on the cleaned data frames defClean, offClean, and bioClean.

1. Reformat the defClean data frame into a tidy format using pivot_longer. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.

2. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?

3.Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense.
```{r}

```